### 2018-11-15 trunk/chess/papers/survival.tex
Footnote 1

@@

Neither side has a decisive advantage, and many game end in a draw,
with both kings surviving. So the survival chances of a king are
clearly greater than 50\%---pretty decent odds. Is it possible that
any other piece has even better chances?

@@

>>>>

If you are forced to play chess to the death, you are in trouble,
because most people are not good at chess (for example, the author)
and yet want to live.\footnote{It is easy for two players to
  collaborate to produce a draw, especially by simply agreeing to a
  draw at the outset of the game (if allowed). Some tournament formats
  forbid the players from agreeing to a draw verbally before a certain
  point (e.g. 30 moves). There are always other routes to a draw, for
  example by stalemate or repeating the same position three times.
  Collaboratively producing such situations is easy, but this strategy
  is not likely a stable equilibrium: Players can often gain a
  substantial advantage by going ``off script'' and instead trying to
  win the game. Additionally, sometimes the terms of
  chess-to-the-death do not allow the players to communicate at all
  beforehand, nor during the game. If this is the case, then it may be
  difficult to agree on the approach to drawing, let alone establish
  that this is both player's desire. Since the rules of
  chess-to-the-death can't forbid us from colluding at this moment, I
  hereby declare that the following is the correct approach: 
%
  1. Nf3. This is a reasonable opening move for white (``Zukertort opening'')
  which can transpose into several common systems (e.g. King's
  Indian). Since the knight can move back to g1 on the next move,
  knight moves are the fastest route to a draw by repetition. This
  has a good chance of signaling to a wise player that a draw is
  desired. The player should make this move after pondering carefully
  for some time, and then looking meaningfully into the other player's
  eyes. 
  1. ... Nf6. This is both a strong response for black in a real game,
  and simultaneously a signal that a draw is desired. The other
  advantage is that very weak players may simply copy what white does.
  In doing so, they will also play this move. 
  2. Ng1?!. This is a terrible move for white, but clearly signals
  the intention to draw.
  2. ... Ng8!. ``Fool's Draw Accepted.'' The starting position is reached
  for a second time.
  3. Nf3 Nf6. At this point the game should clearly continue repeating
  the sequence, although since we are in the start position, white has
  any number of strong opening moves available. Signaling the draw line
  and then 3. d4!? may be pyschologically devastating.
  4. Ng1 Ng8 1/2-1/2. The starting position is reached for the third
  time, which by rule is a draw.

  Since this is one of the shortest possible routes to a draw, I hereby
  dub this line the ``Fool's Draw,'' by analogy with the Fool's Mate.

  If 1. ... Nc6 or another Knight's move, white can also consider
  continuing in the obvious way. However after 1. ... d5, black
  has refused or not noticed the draw. Fortunately, white is still
  in a good position to play the game normally. White can try
  to be more obvious with 2. Ng1, but if black is choosing to just
  play normally, white takes a distinct handicap by doing so.

  The biggest risk for white is that black does not play 2. ... Ng8
  but rather a normal move like 2. ... g6 (``Fool's Draw Betrayed'').
  This can happen if black is not metagaming at all (for Nf6 is a
  normal response to the normal Nf3), or if black is an exceptionally
  shrewd metagamer (tricking white into wasting two tempos with Ng1
  by pretending to be cooperating).

  Of course, this all relies on the assumption that if
  chess-to-the-death ends in a draw, the players are spared or allowed
  to repeat indefinitely. If both players are actually executed, then
  this line is truly a Fool's Draw! }
%
But what if you are forced to {\it be one of the chess pieces} to the
death? That is, your little soul inhabits one of the 32 pieces or
pawns and your soul is vanquished if that piece is eliminated. Now it
doesn't matter whether you're good or bad at chess, but

@@

Neither side has a decisive advantage, and many games end in a draw,
with both kings surviving. So the survival chances of a king are
clearly greater than 50\%---pretty decent odds. Is it possible that
any other piece has even better chances?

@@

% Battle chess
% Feminist game
